Leadership across cultures and contexts: an exploratory analysis

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Working on discussion.

## Clusters of leader qualities

Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019]. Exploratory cluster analyses reported here suggest multiple leadership phenotypes.

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories. Variables were clustered together if they were positively correlated. Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster. See Figure \@ref(fig:cluster)A. Several of these traits often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

The remaining social benefit variables included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005]. A few variables did not seem to fit the social benefit theme, however, such as *Xenophobia* and *Drug consumption* (although drug use is often an important social activity).

@@

## Clusters of leader functions

>>>>

## Features of human leadership

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful groups. Variables were clustered together if they were positively correlated. Leadership and high social status are expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019]. Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies, prestige or dominance; for example, @van_vugt_dual_2019 [p. 1] "argue that that there are two leadership styles: prestige style and dominance style" and @cheng_two_2013 [p. 119] in focusing on dominance-prestige as two distinct leadership strategies suggest that, "that human social stratification is characterized by the co-occurrence of both strategies."

Exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts. Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.

### Clusters of leader qualities

Leader qualities formed two top-level clusters, which we identify as a *Prosocial qualities* cluster, and an *Individual qualities* cluster, each of which contained two interpretable and generally supported subclusters. See Figure \@ref(fig:cluster)A.

The *Prosocial qualities* cluster comprised subclusters we labeled, *Cultural conformity* and *Prosocial competencies*, the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005]. The larger *Prosocial qualities* cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership. A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical fromidable*, and *Drug consumption* (although drug use is often an important social activity).

Several of these traits often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

@@

### Clusters of leader functions
